Run-Time Monitoring with Adjustable Overheads \\Using Dataflow-Guided Filtering

### 2014-09-13 dropping.tex
Cut to 11 pages.
Edits from Tao's comments.

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented in order to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating metadata with array pointers that 
indicates the array's base (start) and bound (end) addresses. On loads or stores with the
array
pointer, the monitor checks that the memory address accessed is within the base and
bound addresses. In addition, this base and bound metadata
is propagated on ALU and memory instructions to track the corresponding array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

@@

Figure~\ref{fig:dropping.example_full} shows an example pseudo-code segment,
its assembly level instructions, and the corresponding monitoring operations
for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc} (line 1). 
As {\tt malloc} returns the array's address in a register, the monitor
associates base and bounds metadata with the corresponding register. 
Next, pointer {\tt y} is set to point to the middle of array {\tt x} (line 2). 
At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
The monitor propagates the metadata of the original pointer in {\tt r1} to the
metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
Line 3a shows setting register {\tt r3} to a constant value of 1.
When this happens,
{\tt r3}'s metadata is also reset to NULL in case it previously stored a
pointer. 
Finally, the value of {\tt r3} is written to memory using both pointers {\tt x} and {\tt y}.
In both cases, the monitor checks whether the store address is within the bounds of the register metadata. In the first case (line 3b), {\tt x+12}
is within the original array's bounds. No error is raised and the metadata of
{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
pointer, than this would allow a future instruction to load the pointer and use
it to access the corresponding array or data structure. In the second case,
{\tt y+12} (line 4) corresponds to {\tt x+20} which is not within the array's bounds and
the monitor will raise an error.

@@

% Dataflow engine high-level
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
    \caption{Hardware support for dropping.}
    \label{fig:dropping.dataflow_overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/dataflow_architecture.pdf}
    \vspace{-0.2in}
    \caption{Hardware architecture of the dataflow engine.}
    \label{fig:dropping.dataflow} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

>>>>

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06, eraser-tocs97}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented in order to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating metadata with array pointers that 
indicates the array's base (start) and bound (end) addresses. On loads or stores with the
array
pointer, the monitor checks that the memory address accessed is within the base and
bound addresses. In addition, this base and bound metadata
is propagated on ALU and memory instructions to track the corresponding array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

@@

Figure~\ref{fig:dropping.example_full} shows an example pseudo-code segment,
its assembly level instructions, and the corresponding monitoring operations
for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc} (line 1). 
As {\tt malloc} returns the array's address in a register, the monitor
associates base and bounds metadata with the corresponding register. 
Next, pointer {\tt y} is set to point to the middle of array {\tt x} (line 2). 
At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
The monitor propagates the metadata of the original pointer in {\tt r1} to the
metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
Line 3a shows setting register {\tt r3} to a constant value of 1.
When this happens,
{\tt r3}'s metadata is also reset to NULL in case it previously stored a
pointer. 
Finally, the value of {\tt r3} is written to memory using both pointers {\tt x} and {\tt y}.
In both cases, the monitor checks whether the store address is within the bounds of the register metadata. In the first case (line 3b), {\tt x+12}
is within the original array's bounds. No error is raised and the metadata of
{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
pointer, then this would allow a future instruction to load the pointer and use
it to access the corresponding array or data structure. In the second case,
{\tt y+12} (line 4) corresponds to {\tt x+20} which is not within the array's bounds and
the monitor will raise an error.

@@

% Dataflow engine high-level
\begin{figure}
  \begin{center}
    \vspace{-0.1in}
    \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
    \vspace{-0.2in}
    \caption{Hardware support for dropping.}
    \label{fig:dropping.dataflow_overview}
    \vspace{-0.2in}
  \end{center}
\end{figure}

@@

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/dataflow_architecture.pdf}
    \vspace{-0.3in}
    \caption{Hardware architecture of the dataflow engine.}
    \label{fig:dropping.dataflow} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

### 2014-09-13 evaluation.tex
Edits from Tao's comments.

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We implement the monitor as a software-based monitor
running on a parallel processor core, similar to LBA \cite{lbaa-asid06}.
We model the main and monitoring cores as
running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.

@@

In this section, we evaluate the effectiveness of
using partial monitoring to trade-off coverage for reduced overheads.
For these results, we use a slack-based, unrestricted dropping policy.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of dynamic checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC and LS). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system, 
assuming that errors/attacks are uniformly distributed across checks. 
%While we could not evaluate real errors/attacks due to difficulty in setting up
%a large number of bugs and exploits, we believe that the percentage of checks
%provides a good estimate of detection probability when errors/attacks are 
%uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

@@

Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
benchmarks, source-only drop fails to meet the specified overhead target.
The overshoot of the overhead target is quite high with an overhead differences
of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
Since only array allocations are considered as source events for BC, it is
difficult for source-only dropping to match overhead targets. Although
Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
source-only dropping, this is largely due to the fact that is running with
higher overheads.

>>>>

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We implement the monitor as a software-based monitor
running on a parallel processor core, similar to LBA \cite{lba-asid06}.
We model the main and monitoring cores as
running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.

@@

In this section, we evaluate the effectiveness of
using partial monitoring to trade-off coverage for reduced overheads.
For these results, we use a slack-based, unrestricted dropping policy.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. The bottom portion of each
bar shows the coverage for a 1.1x overhead target and the additional
coverage for increased overhead targets are stacked above this.
We define \emph{monitoring coverage} as the 
percentage of dynamic checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC and LS). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system, 
assuming that errors/attacks are uniformly distributed across checks. 
%While we could not evaluate real errors/attacks due to difficulty in setting up
%a large number of bugs and exploits, we believe that the percentage of checks
%provides a good estimate of detection probability when errors/attacks are 
%uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

@@

Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
benchmarks, source-only drop fails to meet the specified overhead target.
The overshoot of the overhead target is quite high with an overhead differences
of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
Since only array allocations, which are infrequent, are considered as source events for BC, it is
difficult for source-only dropping to match overhead targets. Although
Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
source-only dropping, this is largely due to the fact that is running with
higher overheads.
